---
title: "STAT341_A2"
author: "Zhi Kai Chen"
date: "2/18/2022"
header-includes: \usepackage[makeroom]{cancel}
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(car)
library(readr)

```
#### Question One - 15 Marks

```{r}
econ_mob_data <- read_csv("EconomicMobility.csv")
```

a) **[4 marks]** Construct a 2X1 figure containing two scatter-plots. The first plot is latitude (y) vs longitude (x) in which the points are shaded/coloured by population categories below.


```{r}

# Create new columns for population and commuting colour. Set Black as default Colour for both
econ_mob_data$Population_Colour="black"
econ_mob_data$Commute_Colour= "black"

# Colour population and commute based on conditions

# Conditions for population
econ_mob_data$Population_Colour[econ_mob_data$Population <= 100000]="blue"
econ_mob_data$Population_Colour[econ_mob_data$Population > 100000 & econ_mob_data$Population < 500000]="red"
econ_mob_data$Population_Colour[econ_mob_data$Population > 500000 & econ_mob_data$Population < 5000000]="green"
econ_mob_data$Population_Colour[econ_mob_data$Population >= 5000000]="black"

# Conditions for commute
econ_mob_data$Commute_Colour[econ_mob_data$Commute <= 0.25]="blue"
econ_mob_data$Commute_Colour[econ_mob_data$Commute > 0.25 & econ_mob_data$Commute < 0.5]="red"
econ_mob_data$Commute_Colour[econ_mob_data$Commute > 0.5 & econ_mob_data$Commute < 0.75]="green"
econ_mob_data$Commute_Colour[econ_mob_data$Commute >= 0.75]="black"


par(mfrow = c(1,2))
# Scatter Plot coloured by Population Categories
plot(Latitude~Longitude, data = econ_mob_data, main = "Population Category", font.main=1, 
     col = econ_mob_data$Population_Colour)
# Scatter Plot coloured by Commuting Category
plot(Latitude~Longitude, data = econ_mob_data, main = "Commute Category", font.main=1, 
     col = econ_mob_data$Commute_Colour)
```

For the Population Category ScatterPlot it seems that there are identifiable clusters. For example you can see that there is a large cluster of blue which means that low population communties are located near each other according to their longitude and latitude.

For the Commute Category ScatterPlot there are also are identifiable clusters. This suggests that people living around the same area (similar longitude and latitude coordinates), commute similar amounts.

There does not seem to be any relationship between population and commute. The cluster colours are not in the same locations

b) **[3 marks]** We are interested in the relationship between the fraction of workers with a commute of less than 15 minutes (y) and the population of the communities (x). 


```{r}
powerfun <- function(x,y, alpha_x, alpha_y){
  if (alpha_x == 0)
    x <- log(x)
  if (alpha_y ==0)
    y <- log(y)
  else if (alpha_x > 0){
    x <- x^alpha_x
  }
  else if(alpha_y > 0){
    y <- y^alpha_y
  }
  else if(alpha_x < 0){
    x <- -x^alpha_x
  }
  else if(alpha_y <0){
    y <- -y^alpha_y
  }
  
  return(list(x_transformed = x, y_transformed = y))
}
transformed <- powerfun(econ_mob_data$Population, econ_mob_data$Commute, -0.5, -0.5)


par(mfrow = c(1,2))
# Original Commute vs Population Scatterplot
plot(Commute~Population, data= econ_mob_data, main = "Commute vs Population", font.main=1)
# Transformed Commute vs Population Scatterplot
plot(transformed$x_transformed, transformed$y_transformed , main = bquote(alpha == -.25 ), font.main=1, xlab = bquote(-("Population"^alpha )),
     ylab = bquote(-("Commute"^alpha )))

```
c) **[5 marks]** Using power-transformations, we would like to straighten the scatterplot in part (b). Suppose the strength of the linear relationship between two variables $X$ and $Y$ is measured by the attribute $L=1-\rho_{_{XY}}^2$ in which $\rho_{_{XY}}$ is the correlation coefficient between the two variables. Assuming $x=population$ and $y=commute$, find the power-transformed data $(T_{\alpha_x}(x),T_{\alpha_y}(y))$ which have the strongest linear relationship measured by $L$. For full marks, you should:

    + Create a function $L$ in `R` which takes in $\theta=(\alpha_x,\alpha_y)$ as input and outputs $1-corr^2\left(T_{\alpha_x}(x) , T_{\alpha_y}(y)\right)$.
    + Optimize the function $L$ with initial values $\alpha_x=\alpha_y=1$, and report optimum values of $\alpha_x$ and $\alpha_y$.
    + Generate the scatterplot of the original data as well as the scatterplot of the power-transformed data, and calculate the correlation coeeficient for each scatterplot
    
    While you have learned the gradient descent method for optimization, it requires deriving the gradient function. The built-in function `nlminb` can use different numerical optimization methods to find the minimum of a function. Read the help document of `nlminb` carefully, and use it for to solve your optimizaiton problem in this question.
