---
title: "SYDE334_A3"
author: "Zhi Kai Chen"
date: "3/06/2022"
header-includes: \usepackage[makeroom]{cancel}
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(car)
library(readr)

```
(a) Fitting model and providing a fit summary
```{r}
insurance_data <- read.table('insurance.txt', sep=",", header = T) 
fit1 = lm(charges~age+sex+bmi+children+smoker+region, data = insurance_data)
summary(fit1)
```
(b) Comment on overall significance of the model

To test the significance $H_0: \beta_1 = \beta_2 = \beta_3 = \beta_4 = \beta_5 = \beta_6 = 0$. We reject $H_0$ since the F statistics takes a value of 500.8 and the p-value is much smaller than 0.001 (p-value < 2.2e-16). The coefficient of determination is 0.7509 indicating that this is a decent fit.

c) Focus on regional difference


i) northeast and northwest

t-test is used to test the hypothesis of the significance of the northwest region (no northeast region coefficient)

$$H_0: \beta_{nw} = 0$$
$$H_A: \beta_{nw} \neq 0$$ 
From the p-value output, the value is 0.458769 which is not less than 0.05 therefore this region is not significant.

ii) southeast and southwest

t-test is used to test the hypothesis of the significance of the southeast and southwest region 

for southeast
$$H_0: \beta_{se} = 0$$
$$H_A: \beta_{se} \neq 0$$ 
From the p-value output, the value is 0.030782 which is less than 0.05 therefore the southeast region is significant.

for southwest 
$$H_0: \beta_{sw} = 0$$
$$H_A: \beta_{sw} \neq 0$$ 
From the p-value output, the value is 0.044765 which is less than 0.05 therefore the southwest region is significant.


iii) northeast and southeast

iv) northwest and southwest



d) Is region an important predictor for health insurance cost?

Some regions display signifance therefore some regions are important predictors for health insurance cost.

The Southwest and Southeast regions display p-value of 0.044765 and 0.030782 respectively. Both these values are less than 0.05 therefore these regions are significant.

The northwest region on the other hand has a p-value of 0.458769. This value is greather than 0.06 therefore this region is not significant.



