---
title: "SYDE334-Assignment-4"
author: "Zhi Kai Chen"
date: '2022-03-27'
output:
  html_document:
    df_print: paged
  pdf_document: default
header-includes: \usepackage[makeroom]{cancel}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(car)
library(readr)
library(plotly)

```

### Question Three

```{r}
hosp_data <- read.table("hospitalstay.dat", header = TRUE)
```

#### a) Fitting Regression Model

```{r}
fit = lm(Length~Culture+Xray+Region+PatNum, data = hosp_data)
summary(fit)
```
Using the summary results from the lm fit. We can see that **Culture** and **Xray** are not important explanatory variables. This is obtained based on t-test for $H_0: \beta_j = 0$.

#### b) Residual Plots for Model Assumptions

```{r}
par(mfrow = c(1,2))

plot(fit$fitted.values, rstudent(fit), xlab = 'Fitted Values', ylab = 'Studentized Residuals')
abline(h=c(2,-2), col = 'red')
plot(hosp_data$Length, rstudent(fit), xlab = 'Length', ylab = 'Studentized Residuals')
```

The Studentized Residuals vs Fitted Values plot show a horizontal band with most of the values falling between 2 and -2. This indicates that there is no visible defects


#### c) Checking for potential outliers and influente observations

```{r}
par(mfrow=c(1,3))

plot(fit$fitted.values, rstudent(fit), xlab = "Fitted Values", ylab = "Studentized Residuals")
abline(h=c(3,-3), col = 'red')
plot(hatvalues(fit), ylab = "Leverage")
abline(h=2*mean(hatvalues(fit)), col = 'red')
plot(cooks.distance(fit), ylab = "Cook's Distance")
abline(h=c(0.5), col = 'red')
```
```{r}
hosp_data[which(rstudent(fit)>3),]
```
Using the studentized residuals, entries 45 and 46 are considered outlier in response.

```{r}
hosp_data[hatvalues(fit)>2*mean(hatvalues(fit)),]
```
Using the leverage, entries 4,7,10,12,45,53,103,109, and 111 are potentially outlier in explanatory variable


```{r}
hosp_data[which(cooks.distance(fit)>0.5),]
```

Using the Cooks Distance, we can see entry 45 is an infuential observation

#### d) Removing influential observations and fitting data
```{r}
fit1 = lm(Length~Culture+Xray+Region+PatNum, data = hosp_data[-45,])
summary(fit1)
```


There is no extreme difference in the estimate of the predictors, however there is a difference in which explanatory variables are significant. Before removing the influential case, **Xray** was not an important variable. Now that we have removed the influential case and fit the model, we can see that **Xray** is a statistically significant variable. **Region** and **PatNum** remain significant, **Culture** remains insignificant.