---
title: "STAT341-Assignment-4"
author: "Zhi Kai Chen"
date: '2022-03-26'
output:
  pdf_document: default
  html_document:
    df_print: paged
header-includes: \usepackage[makeroom]{cancel}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(car)
library(readr)
library(plotly)

```
#### Question Two - 25 Marks

a) **[5 marks]** Generate six scatter plots  of the data in a $3\times2$ grid, where the plots are data as well as polynomials of degrees 1, 2, 5, 10, 15, and 20, respectively overlaid. Use the `getmuhat` function defined in the lectures to estimate these polynomial predictor functions. Use a different colour for each of the different degrees, and use a legend to indicate which degree polynomial is visualized in each plot.

$\;$ 

```{r}
ozone_data <- read_csv("OzoneData1.csv")
```
```{r}
getmuhat <- function(sampleXY, complexity = 1) {
  formula <- paste0("y ~ ", 
                    if (complexity==0) {
                      "1"
                      } else
                        paste0("poly(x, ", complexity, ", raw = FALSE)")
                        #paste0("bs(x, ", complexity, ")")
  )
  
  fit <- lm(as.formula(formula), data = sampleXY)
  tx = sampleXY$x
  ty = fit$fitted.values
  
  range.X = range(tx)
  val.rY = c( mean(ty[tx == range.X[1]]), 
              mean(ty[tx == range.X[2]]) )
  
  ## From this we construct the predictor function
  muhat <- function(x){
    if ("x" %in% names(x)) {
      ## x is a dataframe containing the variate named
      ## by xvarname
      newdata <- x
    } else
      ## x is a vector of values that needs to be a data.frame
    { newdata <- data.frame(x = x) }
    ## The prediction
    ##
    val = predict(fit, newdata = newdata)
    val[newdata$x < range.X[1]] = val.rY[1]
    val[newdata$x > range.X[2]] = val.rY[2]
    val
  }
  ## muhat is the function that we need to calculate values
  ## at any x, so we return this function from getmuhat
  muhat
}

n = 550
x <- 1:550
y <- ozone_data$Ozone

ozone_data_sample <- data.frame(x,y)



muhat_1 <- getmuhat(ozone_data_sample, 1)
muhat_2 <- getmuhat(ozone_data_sample, 2) 
muhat_5 <- getmuhat(ozone_data_sample, 5)
muhat_10 <- getmuhat(ozone_data_sample, 10)
muhat_15 <- getmuhat(ozone_data_sample, 15)
muhat_20 <- getmuhat(ozone_data_sample, 20)

xlim <- extendrange(ozone_data_sample[x,])


par(mfrow=c(3,2), mar=c(2,2,2,2))

plot(ozone_data_sample, pch=19, col = adjustcolor("black", 0.5), xlab="Day", ylab="Ozone")
curve(muhat_1, from = xlim[1], to = xlim[2], add = TRUE, 
      col = "red", lwd=2, main = "Degree = 1")

plot(ozone_data_sample, pch=19, col = adjustcolor("black", 0.5), xlab="Day", ylab="Ozone")
curve(muhat_2, from = xlim[1], to = xlim[2], add = TRUE, 
      col = "blue", lwd=2, main = "Degree = 2")

plot(ozone_data_sample, pch=19, col = adjustcolor("black", 0.5), xlab="Day", ylab="Ozone")
curve(muhat_5, from = xlim[1], to = xlim[2], add = TRUE, 
      col = "green", lwd=2, main = "Degree = 5")

plot(ozone_data_sample, pch=19, col = adjustcolor("black", 0.5), xlab="Day", ylab="Ozone")
curve(muhat_10, from = xlim[1], to = xlim[2], add = TRUE, 
      col = "orange", lwd=2, main = "Degree = 10")

plot(ozone_data_sample, pch=19, col = adjustcolor("black", 0.5), xlab="Day", ylab="Ozone")
curve(muhat_15, from = xlim[1], to = xlim[2], add = TRUE, 
      col = "purple", lwd=2, main = "Degree = 15")

plot(ozone_data_sample, pch=19, col = adjustcolor("black", 0.5), xlab="Day", ylab="Ozone")
curve(muhat_20, from = xlim[1], to = xlim[2], add = TRUE, 
      col = "pink", lwd=2, main = "Degree = 20")


par(fig = c(0, 1, 0, 1), oma = c(0, 0, 0, 0), mar = c(5, 0, 0, 0), new = TRUE)
   plot(0, 0, type = 'l', bty = 'n', xaxt = 'n', yaxt = 'n')

legend(x = "top",
       legend = c("1", "2", "5", "10", "15", "20"),
       lty = c(1,1,1,1,1,1),
       col = c("red", "blue", "green", "orange", "purple", "pink"),
       xpd = TRUE,
       lwd = 5,
       horiz = TRUE,
       cex = 1,
       seg.len = 1,
       bty = 'n')

```
b) **[6 Marks]** generate $M=50$ samples $\mathcal{S}_1,\mathcal{S}_2,\ldots,\mathcal{S}_{50}$ of size $n=100$. You are encouraged (but don't have to) use functions `getSampleComp` and `getXYSample` from the lectures. Fit polynomials of degree 1, 2, 5, 10, 15, and 20 to every sample. Now, create another $3\times2$ grid of plots like in part (a)

```{r}
getSampleComp <- function(pop, size, replace=FALSE) {
  N <- popSize(pop)
  samp <- rep(FALSE, N)
  samp[sample(1:N, size, replace = replace)] <- TRUE
  samp
}

getXYSample <- function(xvarname, yvarname, samp, pop) {
  sampData <- pop[samp, c(xvarname, yvarname)]
  names(sampData) <- c("x", "y")
  sampData
}


popSize <- function(pop) {nrow(as.data.frame(pop))}
sampSize <- function(samp) {popSize(samp)}

N_S <- 50
set.seed(341)
n = 100



samps <- lapply(1:N_S, FUN= function(i){getSampleComp(ozone_data_sample, n)})
Ssamples <- lapply(samps, 
                   FUN= function(Si){getXYSample("x", "y", Si, ozone_data_sample)})
Tsamples <- lapply(samps,
                   FUN= function(Si){getXYSample("x", "y", !Si, ozone_data_sample)})

muhats_1 <- lapply(Ssamples, getmuhat, complexity = 1)
muhats_2 <- lapply(Ssamples, getmuhat, complexity = 2)
muhats_5 <- lapply(Ssamples, getmuhat, complexity = 5)
muhats_10 <- lapply(Ssamples, getmuhat, complexity = 10)
muhats_15 <- lapply(Ssamples, getmuhat, complexity = 15)
muhats_20 <- lapply(Ssamples, getmuhat, complexity = 20)



xlim <- extendrange(ozone_data_sample[x,])


par(mfrow=c(3,2), mar=c(2,2,2,2))

plot(ozone_data_sample, pch=19, col = adjustcolor("black", 0.5), xlab="Day", ylab="Ozone")

for (i in 1:N_S){
  curveFn <- muhats_1[[i]]
  curve(curveFn, from = xlim[1], to = xlim[2], add = TRUE, 
      col = adjustcolor("red", 0.2), lwd=2, lty = (1), main = "Degree = 1")
}

plot(ozone_data_sample, pch=19, col = adjustcolor("black", 0.5), xlab="Day", ylab="Ozone")

for (i in 1:N_S){
  curveFn <- muhats_2[[i]]
  curve(curveFn, from = xlim[1], to = xlim[2], add = TRUE, 
      col = adjustcolor("blue", 0.2), lwd=2, lty = (1), main = "Degree = 2")
}

plot(ozone_data_sample, pch=19, col = adjustcolor("black", 0.5), xlab="Day", ylab="Ozone")

for (i in 1:N_S){
  curveFn <- muhats_5[[i]]
  curve(curveFn, from = xlim[1], to = xlim[2], add = TRUE, 
      col = adjustcolor("green", 0.2), lwd=2, lty = (1), main = "Degree = 5")
}

plot(ozone_data_sample, pch=19, col = adjustcolor("black", 0.5), xlab="Day", ylab="Ozone")

for (i in 1:N_S){
  curveFn <- muhats_10[[i]]
  curve(curveFn, from = xlim[1], to = xlim[2], add = TRUE, 
      col = adjustcolor("orange", 0.2), lwd=2, lty = (1), main = "Degree = 10")
}

plot(ozone_data_sample, pch=19, col = adjustcolor("black", 0.5), xlab="Day", ylab="Ozone")

for (i in 1:N_S){
  curveFn <- muhats_15[[i]]
  curve(curveFn, from = xlim[1], to = xlim[2], add = TRUE, 
      col = adjustcolor("purple", 0.2), lwd=2, lty = (1), main = "Degree = 15")
}

plot(ozone_data_sample, pch=19, col = adjustcolor("black", 0.5), xlab="Day", ylab="Ozone")

for (i in 1:N_S){
  curveFn <- muhats_20[[i]]
  curve(curveFn, from = xlim[1], to = xlim[2], add = TRUE, 
      col = adjustcolor("pink", 0.2), lwd=2, lty = (1), main = "Degree = 20")
}


par(fig = c(0, 1, 0, 1), oma = c(0, 0, 0, 0), mar = c(5, 0, 0, 0), new = TRUE)
   plot(0, 0, type = 'l', bty = 'n', xaxt = 'n', yaxt = 'n')

legend(x = "top",
       legend = c("1", "2", "5", "10", "15", "20"),
       lty = c(1,1,1,1,1,1),
       col = c("red", "blue", "green", "orange", "purple", "pink"),
       xpd = TRUE,
       lwd = 5,
       horiz = TRUE,
       cex = 1,
       seg.len = 1,
       bty = 'n')

```
(c) **[6 points]** Using the $M=50$ samples of size $n=100$ generated in part (b), calculate the APSE (and each of its components) for degrees `0:15`. In particular, print out a table that shows for each degree `apse`, `var_mutilde`, `bias2` and `var_y`.

```{r}
getmubar <- function(muhats) {
  function(x) {
    Ans <- sapply(muhats, FUN=function(muhat){muhat(x)})
    apply(Ans, MARGIN=1, FUN=mean)
  }
}

getmuFun <- function(pop, xvarname, yvarname){
  pop = na.omit(pop[, c(xvarname, yvarname)])

    # rule = 2 means return the nearest y-value when extrapolating, same as above.
  # ties = mean means that repeated x-values have their y-values averaged, as above.
  tauFun = approxfun(pop[,xvarname], pop[,yvarname], rule = 2, ties = mean)
  return(tauFun)
}



apse_all <- function(Ssamples, Tsamples, complexity, tau){
  ## average over the samples S
  ##
  N_S <- length(Ssamples)
  muhats <- lapply(Ssamples, 
                   FUN=function(sample) getmuhat(sample, complexity)
)
## get the average of these, mubar
mubar <- getmubar(muhats)

rowMeans(sapply(1:N_S, 
                FUN=function(j){
                  T_j <- Tsamples[[j]]
                  S_j <- Ssamples[[j]]
                  muhat <- muhats[[j]]
                  ## Take care of any NAs
                  T_j <- na.omit(T_j)
                  y <- c(S_j$y, T_j$y)
                  x <- c(S_j$x, T_j$x)
                  
                  tau_x <- tau(x)
                  muhat_x <- muhat(x)
                  mubar_x <- mubar(x)
                  
                  apse <- (y - muhat_x)
                  bias2 <- (mubar_x - tau_x)
                  var_mutilde <- (muhat_x - mubar_x)
                  var_y <- (y - tau_x)
                  
                  squares <- rbind(apse, var_mutilde, bias2, var_y)^2
                  
                  ## return means
                  rowMeans(squares)
                }
  ))
}



muhat = getmuFun(ozone_data_sample, "x", 'y')


degrees <- 0:15

apse_vals <- sapply(degrees, 
                    FUN = function(degree){ 
                      apse_all(Ssamples, Tsamples, 
                               complexity = degree, tau = muhat)
                    }
)

# Print out the results
t(rbind(degrees, apse=round(apse_vals,5)))

```
(d) **[5 points]** Using your results from part (c) construct a plot whose x-axis is degree and which has four lines: one for `apse`, one for `var_mutilde`, one for `bias2` and one for `var_y`. Specifically, and for interpretability, plot `sqrt(apse)`, `sqrt(var_mutilde)`, `sqrt(bias2)` and `sqrt(var_y)` vs. `degree`. Be sure to distinguish the lines with different colours and a legend. Briefly describe the trends you see in the plot.

```{r}
degree = 0:15
plot(degrees[degree], sqrt(apse_vals[1,degree]),
     xlab = "Degree", type = 'l', ylab = 'Attribute Value', col = 'red', lwd = 3, xlim = c(0,15), ylim = c(0,15))
lines(degrees[degree], sqrt(apse_vals[2,degree]), col = 'blue', lwd = 3)
lines(degrees[degree], sqrt(apse_vals[3,degree]), col = 'green', lwd = 3)
lines(degrees[degree], sqrt(apse_vals[4,degree]), col = 'orange', lwd = 3)

```
\bigskip

Looking at the graph, we can see that they are generally flat except for the tail ends of the graph. For lower degrees (0-2), underfitting occurs. For higher degrees (10-15) overfitting occurs.


(e) **[3 points]** Based on your findings in parts (c) and (d), which degree polynomial has the best predictive accuracy? Construct a scatter plot -- like the ones from (a) and (b) -- but this time create just one plot, and overlay just the polynomial predictor function with the degree you identified as best.  

Based off the results from c) and d), it would appear that polynomial with degree 9 has the best predictive accuracy because it has the lowest apse.

```{r}
plot(ozone_data_sample, pch=19, col = adjustcolor("black", 0.5), xlab="Day", ylab="Ozone")
muhat_9 <- getmuhat(ozone_data_sample, 9)
xlim <- extendrange(ozone_data_sample[x,])


curve(muhat_9, from = xlim[1], to = xlim[2], add = TRUE, 
      col = "blue", lwd=2, main = "Degree = 9")
```