---
title: "SYDE334 Assignment 1"
author: "Zhi Kai Chen"
date: "1/28/2022"
header-includes: \usepackage[makeroom]{cancel}
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(car)
library(readr)

```

## Question 3

a) Make a scatter plot of the data. Does there appear to be a linear relationship between
the two variables?

```{r}
flow_data <- read.delim('flowrate.txt')

pressure = flow_data$PressureDrop
flowrate = flow_data$FlowRate
 
plot(pressure, flowrate, xlab = "Pressure Drop (in inches of water)", 
     ylab = "Flow Rate (in m3/min)", 
     main = "Flow Rate vs Pressure Drop")
```
Yes there appears to be a linear relationship between the two variables


b) Fit a simple linear regression model, provide the equation of the fitted line.

```{r}
# First linear model, with intercept
fit = lm(flowrate~pressure, data = flow_data)
summary(fit)

```
Equation of fitted line:

$$\hat{y_i} = -0.179080 + 0.102342x_i + \epsilon$$
$$\epsilon \sim N(0,0.08652^2)$$


c) Conduct a t test (at a 5% significance level) to determine whether or not there is a
linear association between the two variables. State the hypothesis, decision rule, and
conclusion. What is the p-value of your test?

$$ \alpha = 0.05 $$

Hypothesis:
$$ H_0 : \rho = 0 $$ 

Alternate Hypothesis:
$$ H_A : \rho \neq 0 $$

t ratio:
$$ t = \frac{\hat{\beta_1} - 0}{Se(\hat{\beta_1})} = \frac{0.102342}{0.006303} = 16.23703 $$
$$ |t| = |16.23703| > t_{0.05/2, 25-2} = |16.23703| > 2.069$$
Reject the null hypothesis. We can conclude that there is strong evidence of a linear relationship between the flow rate and the pressure drop.

d) Test the hypothesis (at a 5% significance level) that the average change in the flow rate
associated with a 1 inch increase in the pressure drop is 0.10m3/min, and provide a
conclusion in context.

$$ \alpha = 0.05 $$
$$ H_0 : \hat{\beta_0} = 0.10 $$ 
$$ H_0 : \hat{\beta_0} \neq 0.10 $$ 
$$ t = \frac{\hat{\beta_1} - 0.10}{Se(\hat{\beta_1})} = \frac{0.102342-0.10}{0.006303} = 0.3715691 $$
$$ |t| = |0.3715691| < t_{0.05/2, 25-2} = |0.3715691| < 2.069$$
We do not reject the null hypothesis at a 5% level.

e) Test the hypothesis (at a 5% significance level) that, at x = 0, the average flow rate is
-0.1m3/min, and provide a conclusion in context.

f) Estimate the average flow rate when the pressure drop is 10 inches. Provide a 90%
confidence interval on the mean flow rate.

$$\hat{y_p} = \hat{\beta_0} + \hat{\beta_1}(10) = -0.179080 + 0.102342(10) = 0.84434$$
```{r}
x_bar <- mean(pressure)
y_bar <- mean(flowrate)

sxx <- sum((pressure - x_bar)^2)
sxy <- sum((pressure - x_bar)*(flowrate - y_bar))

b1hat <- sxy/sxx
b0hat <- y_bar-b1hat*x_bar

yhat <- b0hat + b1hat*pressure
r <- flowrate - yhat
MSE <- sum(r^2)/(length(pressure)- 2)

#sxx
sxx
#MSE
MSE
# Mean pressure
x_bar
```
$$ se(\hat{y_p}) = \sqrt{S^2[\frac{1}{n} + \frac{(10 - \overline{x})^2}{S_{xx}}]} = \sqrt{0.007485754^2[\frac{1}{25} + \frac{(10 - 10.628)^2}{188.4504}]}  = 0.001854524557 $$
A 90% Confidence Interval for the mean at x = 10
$$\hat{y_p} \pm t_{0.1/2,23}se(\hat{y_p}) =0.84434 \pm (2.069)(0.001854524557) = (0.8405029887, 0.8481770113)$$

g) Provide a 95% prediction interval for the flow rate when the pressure drop was 10 inches.

$$x_p = 10$$
$$\hat{y_p} = \hat{\beta_0} + \hat{\beta_1}(10) = -0.179080 + 0.102342(10) = 0.84434$$
$$ Se(y_p - \hat{y_p}) = \sqrt{(1+\frac{1}{n}+ \frac{(x_p-\overline{x})^2}{S_{xx}})S^2} = \sqrt{(1+\frac{1}{25}+ \frac{(10-10.628)^2}{188.4504})0.007485754^2} = 0.07641678$$


95% Confidence Interval for $y_p$ is:
$$y_p \pm t_{0.05/2, 23}(0.07641678) = 0.84434 \pm 2.069(0.07641678) = (0.6862336822,1.002446318) $$

h) A flow rate of 1.1m3/min was recently observed from the measurement device when the
pressure drop was 10 inches. Using the previous part discuss briefly why you do, or do
not, believe that the measurement system has changed since the data was collected.

The measurement system has changed since the data was collected. Based off the part g), 1.1m3/min does not fall under the confidence interval therefore it is not a possible value.