---
title: "SYDE334_A2_Q2"
author: "Zhi Kai Chen"
date: "2/10/2022"
header-includes: \usepackage[makeroom]{cancel}
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(car)
library(readr)

```

## Question 2

An investigative study collected 40 observations from the Wabash river at random locationsnear Lafayette. Each observation consisted of a measure of water pH (x) and fish count (y). The researchers are interested in how the acidity of the water affects the number of fish.

a) Complete the following ANOVA table for the regression analysis (the p-value need not be exact).
```{r}
# Create the data frame.
regression <- data.frame(
   Source = c("Regression", "Error", "Total"), 
   SumOfSquares = c("50.30", "60.00 - 53.00 = 9.70" ,"60.00"),
   DegreeOfFreedom = c("1","40-1-1 = 38","40-1 = 39"), 
   MeanSquare = c("50.30/1 = 50.30", "9.70/38 = 0.25526315789", ""),
   Fvalue = c("50.3/0.25526315789 = 197.0517", "", ""),
   Pvalue = c("0", "","")
)

regression
```
b) Give the form of the linear model, state the null and alternative hypothesis for the F-test as well as your conclusion in sentence form.

$H_0: x = 0$ 
$H_a: x \neq 0$

The p-value we obtained is 0, with p-value of 0 we must reject the null hypothesis.

c) What is the estimate of the variance of random error?

$$ \hat{\sigma^2} = \frac{SSE}{n -1}  = \frac{9.70}{39} = 0.2487179$$
d) Use the coefficient of determination to explain what percentage of variation in fish count can be explained by water pH?

$$Coef \space Determination = \frac{SSR}{SST} = \frac{50.30}{60.00} = 0.838332$$
83% of variation in fish count can be explained by water pH.




