---
title: "SYDE334_A2_Q1"
author: "Zhi Kai Chen"
date: "2/10/2022"
header-includes: \usepackage[makeroom]{cancel}
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(car)
library(readr)

```

## Question 1

a) Examine the plot of the data. Does a simple linear regression model seem appropriate? Explain

```{r}
stock_data <- read.delim('stock.txt') 
x_vals = stock_data$PortfolioRate
y_vals = stock_data$StockRate
plot(x_vals, y_vals, xlab = "Rate of Return of Market Portfolio", ylab = "Rate of Return of Stock", main = "Stock")

```
Yes, a simple linear regression model seems appropriate. There appears to be a general trend where as the rate of return of market porfolio increases, so does the rate of return of stock.

b) Fit the simple linear regression model, provide the equation of the fitted line. What financial implication does the value of the estimate $\beta_1$ have?

```{r}
# First linear model, with intercept
fit_1 = lm(StockRate~PortfolioRate, data = stock_data) 
summary(fit_1)
```
The financial implication of $\beta_1$ implies that as the rate of return of the portfolio increases by 1, the rate of return of the stock increases by 0.52397.

c) Produce a 95% confidence interval for $\beta_1$, and use it to assess the significance of $\beta_1$.

$$\alpha = 0.05, \space DF = 38$$
$$t_{\alpha/2,38} = t_{0.025,38} = 2.021$$

$se(\hat{\beta_1}) = 0.07461$
$\hat{\sigma}^2 = 2.145$

$$t = \frac{\hat{\beta_1} - \beta_1}{se(\hat{\beta_1})}= \frac{0.52397}{2.145} = 0.244275$$
95% Confidence Interval:

$$\hat{\beta_1} - t_{0.025,n-2}se(\hat{\beta_1}) < \beta_1 < \hat{\beta_1} + t_{0.025,n-2}se(\hat{\beta_1}) $$
$$0.52397 - 2.021*2.145 < \beta_1 < 0.52397 + 2.021*2.145 $$
$$-3.811075 < \beta_1 < 4.859015 $$

d) Test the hypothesis (at a 5% significance level) that is particular stock is an average-risk stock.

e) Predict the return rate for this stock and construct a 95% confidence interval.

```{r}
newdata = data.frame(PortfolioRate = 8)
pred = predict.lm(fit_1, newdata, interval = "confidence", level = 0.95)
pred
```
